---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setopetions,echo=TRUE}
 
```


## Loading and preprocessing the data


```{r preprocessing}
   data<-read.csv("activity.csv",stringsAsFactors=FALSE)
   factoreddata<-transform(data,date=as.Date(date,format="%Y-%m-%d"))
   nonMissingData<-subset(data,data[,"steps"]!="NA")
   dataperday<-aggregate(nonMissingData$steps,by=list(nonMissingData$date),sum) 
   names(dataperday)<-c("date","stepsperday")   
```


## What is mean total number of steps taken per day?

1. Historgram of Number of Steps taken per day

```{r histogram}
   hist(dataperday$stepsperday,main='Histogram of Number of Steps Taken Per Day',xlab='Steps Per Day',col='red')   

```

2. Mean and Median total number of steps taken per day 

```{r meanmedian,results='hide'}
   meanperday<-mean(dataperday$stepsperday)
   medianperday<-median(dataperday$stepsperday)
```

### The Mean steps taken per day are `r meanperday` and Median steps taken per day are `r medianperday`


## What is the average daily activity pattern?

1. Time Series Plot of 5 minute interval and the average number of steps taken

```{r activitypattern}
   steps<-nonMissingData$steps
   plot.ts(x=ts(steps,start=c(0,5),frequency=288),type="l",
           ylab="Average Number of Steps Taken",
           xlab="Time on the basis of 5 Minutes Interval",yaxt="n")
   avg<-signif(mean(steps),digits=4)
   axis(2,at=mean(steps),labels=paste("Average Value",avg,sep="\n"))

```

2. The Five Minute Interval contains Maximum number of steps on average across all the days in the dataset

```{r interval,results='hide'}
   maxStepRow<-subset(nonMissingData,
                      nonMissingData[,"steps"]==max(nonMissingData$steps))
```

### The 5 Minute interval Interval `r maxStepRow$interval-5` to `r maxStepRow$interval` on `r maxStepRow$date` has maximum value of number of steps taken which is `r maxStepRow$steps` 

## Imputing missing values

1. Total number of Missing Values

```{r missing}
   wholerowmissing<-subset(data,is.na(data[,"steps"])|
                              is.na(data[,"date"])|
                              is.na(data[,"interval"]))
   missingCount<-length(wholerowmissing$steps)
   
```

### Total Number of missing rows in this data set is `r missingCount`




## Are there differences in activity patterns between weekdays and weekends?
